### 2009-03-02 doc/storage_system/arc-storage-documentation.tex
updating the storage documentation

@@

@@

The Librarian also accepts report messages from Shepherds, and stores the contents of these messages in the A-Hash. There is one A-Hash object which contains the ID and the expected time of the next heartbeat of the registered Shepherds, and there is one A-Hash object for each Shepherd which stores the states and GUIDs of the files on the Shepherd's storage node. The Librarian checks periodically if there is a Shepherd which is late with its heartbeat, and it changes the states of all the replicas if needed.

@@

@@

A Shepherd service is always connected to a file transfer service (e.g.~`HTTP(S)', `FTP(S)', `ByteIO', `GridFTP', etc.). For each supported file transfer service we need a backend module which makes the Shepherd capable of communicating with the file transfer service to initiate file transfers, to detect whether a transfer was successful or not, to generate local IDs and checksums, etc.

A file in a storage node could be identified with a \emph{referenceID} which is unique within that node. If we know the \emph{location} of a file, which is the ID of the Shepherd service (\emph{serviceID}) and the \emph{referenceID}, we could get the endpoint reference (URL) of the Shepherd from the information system, then we could call its \emph{get} method with the \emph{referenceID} and a list of transfer protocols we can use (e.g.~`HTTP', `FTP'), the Shepherd chooses a protocol from this list which it can provide, and create a transfer URL (\emph{TURL}) and returns it along with the \emph{checksum} of the file. We could download the file from this \emph{TURL}, and verify it with the \emph{checksum}. An end user of the storage system does not need to call this \emph{get} method, because the Bartender service will do it, the user just asks the Bartender and gets the TURL.

@@

With the \emph{stat} method we can get some information about a replica, e.g.~checksum, GUID, state (`creating', `alive' or `invalid'), etc. The \emph{delete} method removes the file.

@@

@@

The current implementation of the Shepherd service has a working \emph{get}, \emph{put}, \emph{stat}, \emph{delete} methods, and a method called \emph{toggleReport} which can be used the simulate storage node failure with the Shepherd not reporting to a Librarian. 
There is a separate service which provide a subset of the ByteIO interface, and there is an other separate service which is a basic HTTP server, these are both could be used as file transfer services, both have its backend module for the Shepherd. Currently both file transfer services have the problem of using too much memory while transferring files.
Further plans include better file transfer services and backend modules for third-party file transfer services.

% subsection prototype_status_and_plans (end)

A file of a Shepherd service is referenced by its \emph{referenceID}. Each file has a \emph{state} which could be `\textbf{creating}' when the transfer is initiated but the file is not uploaded yet, `\textbf{alive}' if the file is uploaded and has a proper checksum, or `\textbf{invalid}' if it does not exists anymore or has a bad checksum. Each file has a \emph{localID} which is used in the backend modules.

@@

\begin{description}
    \item[get(getRequestList)] returns list of (\emph{requestID}, \emph{getResponseData})

    \emph{getRequestList} is a list of (\emph{requestID}, \emph{getRequestData}) where \emph{requestID} is an arbitrary ID used in the reply

    \emph{getRequestData} is a list of (\emph{property}, \emph{value}) pairs, where mandatory properties are: `\textbf{referenceID}' which refers to the file to get and `\textbf{protocol}' indicates a protocol the client can use (there could be multiple protocols in getRequestData).

    \emph{getResponseData} is a list of (\emph{property}, \emph{value}) pairs, such as: `\textbf{TURL}' is a transfer URL which can be used by the client to download the file; `\textbf{protocol}' is the protocol of the TURL; `\textbf{checksum}' is the checksum of the replica; `\textbf{checksumType}' is the name of the checksum method and `\textbf{error}' could contain an error message if there is one.

    \item[put(putRequestList)] returns a list of (\emph{requestID}, \emph{putResponseData})

    \emph{putRequestList} is a list of (\emph{requestID}, \emph{putRequestData}) where \emph{requestID} is an ID used for the response

    \emph{putRequestData} is a list of (\emph{property}, \emph{value}) pairs such as `\textbf{GUID}', `\textbf{checksum}', `\textbf{checksumType}', `\textbf{size}' (the size of the file in bytes), `\textbf{protocol}' (a protocol the client can use, can be multiple) and `\textbf{acl}' (for additional access policy).

    \emph{putResponseData} is a list of (\emph{property}, \emph{value}) pairs such as: `\textbf{TURL}' is the transfer URL where the client can upload the file, `\textbf{protocol}' is the chosen protocol of the TURL and `\textbf{referenceID}' is the generated ID for this new replica, `\textbf{error}' could contain an error message.

    \item[delete(deleteRequestList)] returns a list of (\emph{requestID}, \emph{status})

    \emph{deleteRequestList} is a list of (\emph{requestID}, \emph{referenceID}) pairs selecting the files to remove. The status could be `\textbf{deleted}' or `\textbf{nosuchfile}'
    .
    \item[stat(statRequestList)] returns a list of (\emph{requestID}, \emph{referenceID}, \emph{state}, \emph{checksumType}, \emph{checksum}, \emph{acl}, \emph{size}, \emph{GUID}, \emph{localID})

    \emph{statRequestList} is a list of (\emph{requestID}, \emph{referenceID}) where \emph{referenceID} points to the file whose data we want to get. The method returns all the data the Shepherd know about the replica.

@@

\subsection{Backend modules} % (fold)
\label{sub:backend_modules}

The Shepherd could communicate with the file transfer services via backend modules. Currently there are two kinds of backend modules, one for the \emph{byteio} service (which is a simple implementation of a subset of the ByteIO interface) and one for the \emph{Hopi} service (which is a simple HED-based HTTP server).

@@

In case of the byteio service, there is some small differences. The byteio service does not removes the files from the transfer directory, but it calls the backend module via SOAP to notify it that something is happened. The byteio backend has one SOAP method called `\textbf{notify}':
\begin{description}
    \item[notify(subject, state)] returns `\textbf{OK}' in \emph{notifyResponse}.
\end{description}
When this method is called, the byteio backend module notifies the Shepherd that the file is arrived.

All the backend modules should have this common interface which the Shepherd can use to communicate with the file transfer service:
\begin{description}
    \item[prepareToGet(referenceID, localID, protocol)] returns the \emph{TURL}.

@@

    \item[list()] returns a list of \emph{localID}s currently in the store directory.
    \item[getAvailableSpace()] returns the available disk space in bytes.
    \item[generateLocalID()] returns a new unique \emph{localID}.
    \item[matchProtocols(protocols)] only leave that protocols in the list \emph{protocols} which are supported by this file transfer service.
    \item[checksum(localID, checksumType)] returns the checksum of the file referenced by \emph{localID}, which checksum is generated by the method \emph{checksumType}.

% subsection backend_modules (end)

>>>>

For inter-service authorization we can configure the services to only accept connections from trusted services. We trust a service if we know its DN. For each service we can write trusted DNs in the configuration, or we can periodically get the list from an A-Hash. This can be configured with a `\emph{TrustManager}' section in the service configuration, which contains these entries:

\begin{description}
    \item[DN] is a trusted DN
    \item[CA] is a DN of a trusted CA: we will trust all certificates issued by this CA
    \item[DNsFromAHash] is a list of A-Hash URLs (in \textbf{AHashURL}) from where we periodically get a list of trusted DNs (there can be multiple A-Hash URLs listed here in case of one is offline). This has some attributes: \textbf{CheckingInterval} specifies how frequently we should get the list from the A-Hash (default: 600); \textbf{ID} specifies the ID of the A-Hash objects, where the list is stored (default: `3').
\end{description}

The TrustManager section itself could have some attributes: \textbf{FromFile} could have a filename, if we want to get the whole TrustManager configuration from a separate file; and \textbf{Force} indicates (with `yes' or `no') that we want to ensure that our service does not even get the request if the client of the incoming connection is not trusted.

Example TrustManager configuration:

\begin{verbatim}
    <TrustManager Force="yes">
        <DN>/DC=eu/DC=KnowARC/O=NIIFI/CN=host/epsilon</DN>
        <DN>/DC=eu/DC=KnowARC/O=NIIFI/CN=host/phi</DN>
        <CA>/DC=eu/DC=KnowARC/CN=storage-1233659377.11</CA>
        <DNsFromAHash CheckingInterval="10">
            <AHashURL>https://localhost:60000/AHash</AHashURL>
        </DNsFromAHash>
    </TrustManager>
\end{verbatim}

@@

This can be put in a seperate file, and then use the \textbf{FromFile} attribute of the ClientSSLConfig section to specify where the file is.

@@

The Librarian also accepts report messages from Shepherds, and stores the contents of these messages in the A-Hash. There is one A-Hash object which contains the ID and the expected time of the next heartbeat of the registered Shepherds, and there is one A-Hash object for each Shepherd which stores the states and GUIDs of the files on the Shepherd's storage node (see Section~\ref{sub:shepherds} for details). The Librarian checks periodically if there is a Shepherd which is late with its heartbeat, and it changes the states of all the replicas if needed.

@@

The Librarian has these configuration variables:

\begin{description}
    \item[AHashURL] is the URL of an A-Hash which the Librarian should use. There could be multiple URLs specified here.
    \item[HeartbeatTimeout] specifies in seconds how frequently the Sheperds should send reports to the Librarian.
    \item[CheckPeriod] specifies in seconds how frequently the Librarian should check for late heartbeats.
\end{description}

\begin{verbatim}
    <Service name="pythonservice" id="librarian">
        <ClassName>storage.librarian.librarian.LibrarianService</ClassName>
        <AHashURL>https://localhost:60000/AHash</AHashURL>
        <HeartbeatTimeout>30</HeartbeatTimeout>
        <CheckPeriod>20</CheckPeriod>
        <ClientSSLConfig FromFile="clientsslconfig.xml"/>
        <TrustManager FromFile='librarian_trust_manager.xml'/>
    </Service>
\end{verbatim}

@@

A Shepherd service is always connected to a storage element service (e.g.~a web server). For each supported storage element service we need a backend module which makes the Shepherd capable of communicating with it to initiate file transfers, to detect whether a transfer was successful or not, to generate local IDs and checksums, etc.

A file in a storage node could be identified with a \emph{referenceID} which is unique within that node. If we know the \emph{location} of a file, which is the ID of the Shepherd service (\emph{serviceID}, currently the URL of the Shepherd) and the \emph{referenceID}, we could call the Shepherds \emph{get} method with the \emph{referenceID} and a list of transfer protocols we can use, the Shepherd chooses a protocol from this list which it can provide, and create a transfer URL (\emph{TURL}) and returns it along with the \emph{checksum} of the file. We could download the file from this \emph{TURL}, and verify it with the \emph{checksum}. An end user of the storage system does not need to call this \emph{get} method, because the Bartender service will do it, the user just asks the Bartender and gets the TURL.

@@

With the \emph{stat} method we can get some information about a replica, e.g.~checksum, GUID, state, etc. The \emph{delete} method removes the replica.

@@

The replicas have a state, which could be `\textbf{creating}' when the transfer is initiated but the file is not uploaded yet, `\textbf{alive}' if the file is uploaded and has a proper checksum, or `\textbf{invalid}' if it does not exists anymore or has a bad checksum.

@@

@@

    \item[get(getRequestList)] initiates a download and returns the TURL

    The \emph{getRequestList} is a list of (\emph{requestID}, \emph{getRequestData}) where \emph{requestID} is an arbitrary ID used in the reply, \emph{getRequestData} is a list of (\emph{property}, \emph{value}) pairs, where mandatory properties are: `\textbf{referenceID}' which refers to the file to get and `\textbf{protocol}' indicates a protocol the client can use (there could be multiple protocols in getRequestData).

    The response is a list of (\emph{requestID}, \emph{getResponseData}), where \emph{getResponseData} is a list of (\emph{property}, \emph{value}) pairs, such as: `\textbf{TURL}' is a transfer URL which can be used by the client to download the file; `\textbf{protocol}' is the protocol of the TURL; `\textbf{checksum}' is the checksum of the replica; `\textbf{checksumType}' is the name of the checksum method and `\textbf{error}' could contain an error message if there is one.

    \item[put(putRequestList)] initiates an upload and returns the TURL

    The \emph{putRequestList} is a list of (\emph{requestID}, \emph{putRequestData}) where \emph{requestID} is an ID used for the response,     \emph{putRequestData} is a list of (\emph{property}, \emph{value}) pairs such as `\textbf{GUID}', `\textbf{checksum}', `\textbf{checksumType}', `\textbf{size}' (the size of the file in bytes), `\textbf{protocol}' (a protocol the client can use, can be multiple) and `\textbf{acl}' (which is currently not used).

    The response is a list of (\emph{requestID}, \emph{putResponseData}), where \emph{putResponseData} is a list of (\emph{property}, \emph{value}) pairs such as: `\textbf{TURL}' is the transfer URL where the client can upload the file, `\textbf{protocol}' is the chosen protocol of the TURL and `\textbf{referenceID}' is the generated ID for this new replica, `\textbf{error}' could contain an error message.

    \item[delete(deleteRequestList)] removes a replica.

    The \emph{deleteRequestList} is a list of (\emph{requestID}, \emph{referenceID}) pairs containing the IDs of the files to remove.

    The response is a list of (\emph{requestID}, \emph{status}), where \emph{status} could be `\textbf{deleted}' or `\textbf{nosuchfile}'.
    .
    \item[stat(statRequestList)] returns information about replicas.

    The \emph{statRequestList} is a list of (\emph{requestID}, \emph{referenceID}) where \emph{referenceID} points to the file whose data we want to get.

    The response is a list of (\emph{requestID}, \emph{referenceID}, \emph{state}, \emph{checksumType}, \emph{checksum}, \emph{acl}, \emph{size}, \emph{GUID}, \emph{localID})    
\end{description}

@@

The Shepherd is communicating with the storage element services via backend modules. Currently there are two backend modules implemented, one for the \emph{Hopi} service (which is a simple HED-based HTTP server), one for the \emph{apache} webserver.

@@

All the backend modules should have this common interface which the Shepherd can use to communicate with the storage element services:

\begin{description}
    \item[prepareToGet(referenceID, localID, protocol)] returns the \emph{TURL}.

@@

    \item[list()] returns a list of \emph{localID}s currently in the store directory.
    \item[getAvailableSpace()] returns the available disk space in bytes.
    \item[generateLocalID()] returns a new unique \emph{localID}.
    \item[matchProtocols(protocols)] only leave that protocols in the list \emph{protocols} which are supported by this file transfer service.
    \item[checksum(localID, checksumType)] generates a \emph{checksumType} checksum of the file referenced by \emph{localID}.
\end{description}

The Shepherd has these configuration variables:

\begin{verbatim}
    <Service name="pythonservice" id="shepherd">
        <ClassName>storage.shepherd.shepherd.ShepherdService</ClassName>
        <ServiceID>https://localhost:60000/Shepherd</ServiceID>
        <CheckPeriod>20</CheckPeriod>
        <MinCheckInterval>0.1</MinCheckInterval>
        <StoreClass>storage.store.cachedpicklestore.CachedPickleStore</StoreClass>
        <StoreCfg>
            <DataDir>./shepherd_data1</DataDir>
        </StoreCfg>
        <BackendClass>storage.shepherd.hardlinkingbackend.HopiBackend</BackendClass>
        <BackendCfg>
            <DataDir>./shepherd_store</DataDir>
            <TransferDir>./shepherd_transfer</TransferDir>
            <TURLPrefix>https://localhost:60000/hopi/</TURLPrefix>
        </BackendCfg>
        <LibrarianURL>https://localhost:60000/Librarian</LibrarianURL>
        <BartenderURL>https://localhost:60000/Bartender</BartenderURL>
        <ClientSSLConfig FromFile="clientsslconfig.xml"/>
    </Service>    
\end{verbatim}
